Cultural Evolution of Conservation Behavior at the Intersection of Working

### 2020-10-06 DissertationProposal/DissertationProposal.Rmd
Done before committee review I guess....

# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common however, for these areas to fail to reach their stated conservation goals. The most chronic causes of conservation project failure are the human dimensions. Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors. Research examining how and why conservation practices are adopted has been quite limited; suffering from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

1. Develop a dynamic model to generate predictions regarding if and how conservation norms can emerge and persist via competition between groups.
2. Empirically test theoretical predictions of group conservation adoption as a function of group competition and individual costs and benefits. 
3. Using the ongoing spread of an improved efficiency cookstove design, specifically examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.

@@

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations [@adams_biodiversity_2004; @leverington_global_2010]. While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004]. Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].

Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group [@rogers_diffusion_2010; @simon_win-win_2012]. These behaviors are characterized as individually costly, yet beneficial at the group level. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017]. This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution]. What is less understood however, is the emergence and initial adoption of conservation behaviors.

Research on the diffusion of conservation behaviors suffers from a lack of generalizability. Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors. Excitingly, parallel to the development of conservation social science, the field of evolutionary anthropology has developed the theory of cultural group selection as a mechanistic explanation for the adoption of individually costly, group beneficial behaviors.

@@

Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004]. While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005]. Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

@@

# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
## Background
Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) [@iucn2020protectedplanet]. During this time period, the stated mission of most protected areas has shifted as well. Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities [@fisher_poverty_2007; @naughton-treves_role_2005]. With this recognition, their objectives now commonly include community development goals alongside conservation targets.

```{r paSize, echo=FALSE, fig.cap="Size of newly established no-take protected areas reported by the IUCN 2020", out.width = "100%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/Size.png")
```

Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss [@tallis_ecosystem_2008]. These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations [@hall_tourism_2004; @langholz_perils_2001]. Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not [@fisher_poverty_2007]. Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation [@adams_biodiversity_2004].

Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals [@christie_marine_2004; @watson_performance_2014]. In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands [@clark_protected_2013]. It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations [@bennett_why_2014; @christie_observed_2005].

The degree of local buy-in for protected areas is not random. The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations. A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters [@mills_how_2019]. These findings are particularly worrying given a global decline in government support for land conservation [@watson_performance_2014].

With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research [@bennett_conservation_2017]. This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes. As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups. Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives [@brooks_applying_2018].

Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985]. Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge [@richerson_cultural_2016].

While the theory and principles of cultural group selection are relatively well established, they are yet to be tested mechanistically in the context of protected area adoption. With this in mind, this project aims to demonstrate how the presence of multiple competing groups can facilitate the adoption of adherence to protected areas. In this chapter I will develop a dynamic, agent-based model to test the logic of how protected area norms can emerge under very simple sets of rules in the presence of multiple competing groups, but fail to emerge for solitary groups under identical environmental contexts.

### Question:
Can between group competition provoke protected area adoption where solitary groups may fail to adopt?

### Predictions:
* P1: Adoption of protected area adherence will only emerge under the presence of multiple competing groups.

* P2: Where it already exists, protected area adherence will perpetuate in populations even in solitary groups given group conformity.

* P3: Greater degrees of resource mobility will increase the frequency in which the adoption of protected area adherence emerges in the community(ies).

@@

An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual (Railsback & Grimm, 2012; Wilensky & Rand, 2015). This design mirrors our real-world understanding of the important system components.

### Model Description
In the most simple scenario, the agent based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape. This landscape is divided into two types, working and protected. Resources move between these patches at each time step with some probability. At time = 0, the protected and working landscapes are both at an equal, designated percent carrying capacity. At each time step, agents aim to maximize their harvest up until a designated harvest max. At time = 0, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.

The model also allows for ecological variation during the initial specification. The variable ecological starting conditions are as follows: Resource mobility designates the probability that a resource unit will move from the protected landscape into the working landscape or vice versa at each time step. Protected area size demarcates what percentage of the available resource units are within the protected area. Percent carrying capacity sets the total number of harvestable resource units as a percent of the landscape carrying capacity. Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step.

Agents behave according to a specific set of rules. Agents “choose” to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (defecting). Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e. they do not treat the protected area any differently than the working landscape). Agents learn these behaviors according to their specific learning mechanism. The learning will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents. Behaviors may be modified at each time step by the key processes described below.

For the given number of timesteps, the model follows a set of key processes. At time step one, all agents harvest first spend their initial assigned percentage of effort harvesting from the working landscape. If agents reach harvest max, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect. If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area. They will harvest up to the harvest max if possible. The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape. Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility. If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies. If one agent preformed worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them. If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group. This process repeats for all time steps. This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).

In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above. However, at the end of each time step the groups will compare the average harvest rate each group. Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes. A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group.

In both single and competing group scenarios, the dependent variables of interest are as follows.: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step. Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected. Protected area degradation measures the percent carrying capacity of the protected landscape. Working landscape degradation measures the percent carrying capacity of the working landscape. Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.

@@

@@

## Preliminary Work
I have developed a model which follows the above processes for one group with one resource pool. An interactive application which performs one run of that model can is found [at this link](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).

@@

This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly. Given the theoretical capacity for cultural group selection to provide solutions to pressing environmental issues of collective action, these predictions must be validated experimentally so that they can be implemented effectively.

@@

### Question: 
Are predictions of costly conservation adoption derived from cultural group selection theory significant predictors of real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?

@@

* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such nonprofit conservation initiatives.

## Research Design
Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level. These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs). Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are more opaque. For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.

@@

Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity. Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008]. This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.

@@

### Question
What effect do geographic distance, genetic relatedness, and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?

### Predictions
* P1: Using geographic distance, genetic relatedness, and social network distance as proxies for access to behavioral demonstrators, we expect transmission fidelity to increase with increasing access.

* P2: As access to behavioral demonstration decreases, we expect to see differential changes in the fidelity of replicated behavior and the behavioral performance target of interest.

* P3: Given the high degree of environmental influence on a behavior, we expect to observe a greater change in the fidelity of transmission than in the behavioral performance target, correlative with decreasing access to behavioral demonstrators.

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

@@

The survey instrument used for this research will elicit information regarding the relationship between the builder of each cookstove and the individual they copied their stove design from. Specifically, this survey tool will be used to gather data on geographic distance, genetic relatedness, and social relatedness. I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.

While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics such as the exterior dimensions and size of the openings. I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from. To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions]. Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.

The key research questions for this project will be assessed using multilevel regression operationalized in a Bayesian context. I will compare standardized parameter estimates, to determine which factors (genetic relatedness, geographic distance, social relatedness) most strongly contribute to copying fidelity in this study system. Similar analyses will be used to determine which factors affect cookstove primary outcome variables (fuel efficiency and particulate matter production).

@@

## Preliminary Work
In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:

@@

This dissertation will achieve this goal in three progressive steps. The first will formalize theory around cultural group selection in conservation contexts. This step is critical as conservation is a fundamentally different system from where these theories were originally developed. The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania. Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities, to specifically examine how behavioral efficiency may be altered by social contexts (e.g. communication frequency, genetic relatedness) as a behavior spreads throughout a population. All together, this research will allow us to better understand and possibly predict the emergence and effective spread of conservation norms within a population.

>>>>

# Summary
Area-based conservation is critical for stemming global biodiversity loss. As such, each year billions of dollars are spent to protect key areas. It is common, however, for these areas to fail to reach their stated conservation goals. The most chronic causes of conservation project failure are the human dimensions. Conversely, historic conservation successes are also attributed to widespread human adoption of conservation behaviors. Research examining how and why conservation practices are adopted tends to suffer from a bias toward the reporting of successful projects and highly context dependent drivers and outcomes. This dissertation seeks to demonstrate and better understand the underlying mechanisms which underpin social adoption of conservation practices in and around protected areas. Toward this aim, this project will combine insights from cultural group selection theory with empirical research on Pemba Island, Tanzania to achieve three primary objectives.

1. Develop a dynamic model to generate predictions regarding if and how conservation norms can emerge and persist via competition between groups.

2. Empirically test theoretical predictions of group conservation adoption as a function of group competition and individual costs and benefits.

3. Using the ongoing spread of an improved efficiency cookstove design, specifically examine how the nature of interpersonal relationships may modify conservation behaviors as they spread within groups.

@@

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010]. While high-profile protected area successes are salient in the conservation narrative, protected areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is somewhat speculative as unsuccessful protected areas go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004]. Regardless of the high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].

Community norms around protected area adherence are characterized as individually costly, yet beneficial at the group level. Substantial variation exists in the degree to which communities adhere to protected area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@rogers_diffusion_2010; @simon_win-win_2012]. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017]. This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution]. What is less understood however, is the emergence and initial adoption of conservation behaviors.

Research on the diffusion of conservation behaviors suffers from a lack of generalizability. Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors. Outside of conservation science, the field of evolutionary anthropology has devoted considerable effort to understanding how and why individuals adopt costly cooperative behaviors. Of particular relevance to understanding group adherence to protected area regulations, the theory of cultural group selection offers a mechanistic explanation for the emergence and spread of individually costly group norms.

@@

## Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004]. While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005]. Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

@@

# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
## Background
Cultural group selection theorizes that individuals will adopt cooperative behavior if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985]. Under this framework, the presence of multiple competing groups drives community cooperative norms to emerge under a much broader set of conditions compared to solitary groups [@richerson_cultural_2016]. Protected area conservation is a cooperative phenomenon where individuals sacrifice some degree of their own harvest potential in order to benefit the group as a whole. Therefore, insights from cultural group selection may help conservation scientists better understand and predict community adherence to protected area regulations [@brooks_applying_2018].

While cultural group selection theory offers mechanistic insights for the emergence of cooperative norms in groups broadly, its applicability to conservation specifically is unknown. Conservation has many aspects which make it unique compared to cooperation in the abstract. In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making. Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as mobility and regeneration rate. These specifics make conservation a fundamentally different system from general cooperation, thus we must extend cultural group selection theory considering conservation explicitly.

This project aims to logically test the applicability of cultural group selection to explain the emergence and spread of protected area adoption. In this chapter I will develop a dynamic, agent-based simulation of individuals harvesting resources from a landscape under a variety of environmental constraints. I will then use this simulation to test under what circumstances adherence to protected area regulations emerges as a group norm within solitary and multiple, competing groups. This model will generate qualitative predictions about system outcomes given a very simple set of assumptions.

### Question:
Under what environmental contexts might the mechanisms of cultural group selection influence individual decision making to generate lasting group norms around the adherence to protected area regulations?

### Predictions:
* P1: Groups will adopt protected area regulations in both the solitary and multiple competing groups scenarios under specific environmental contexts.

* P2: In the presence of multiple competing groups, the environmental contexts which generate group adoption of protected area regulations will be much more broad compared to solitary groups.

@@

An agent-based modeling framework is ideal for demonstrating the theory of cultural group selection regarding protected area norms, because it allows such group level properties to emerge out of simple decision rules for each individual [@railsback_agent-based_2012; @wilensky_introduction_2015]. This design mirrors our real-world understanding of the important system components.

### Model Description
In the most simple scenario, the agent based model represents a simplistic world where there is one homogeneous group of agents who harvest resources from a landscape. This landscape is divided into two types, working and protected. Resources move between these patches at each time step with some probability. During initialization, the protected and working landscapes are both at an equal, designated percent carrying capacity. At each time step, agents aim to maximize their harvest up until a designated harvest max. During initialization, agents are assigned as either obligatory cooperators or as having the ability to defect to a small degree, spending a small portion of their time harvesting from the protected landscape.

The model also allows for ecological variation during the initial specification. The variable ecological starting conditions are as follows: Resource mobility designates the probability that a resource unit will move from the protected landscape into the working landscape or vice versa at each time step. Protected area size demarcates what percentage of the available resource units are within the protected area. Percent carrying capacity sets the total number of harvestable resource units as a percent of the landscape carrying capacity. Lastly, the resource regeneration rate designates what percentage the resource regenerates at each time step (see Table \@ref(tab:tabAbmVars)).

@@

Agents behave according to a specific set of rules. Agents choose to spend a given percent of their time harvesting from the protected area (defecting) and a percent of their time harvesting from the working landscape (cooperating). Full defection is when the average percent of time agents spend in the protected area is equal to the percent of the landscape protected (i.e. they do not treat the protected area any differently than the working landscape).

Agents learn their behaviors according to one of two possible learning mechanisms. Agents who learn via success-biased learning will preferentially copy the behavior of more successful agents when they encounter them. Agents who learn via conformist-biased learning will disproportionately copy the behavior of other agents they see as being in the majority, regardless of the success of those agents. The learning strategy of the overall population will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents. Behaviors may be modified at each time step by the key processes described below.

For the given number of timesteps, the model follows a set of key processes. At time step one, all agents first spend their initial assigned percentage of effort harvesting from the working landscape. If agents reach the max harvest amount, they do not harvest from the protected landscape, even if they are assigned as having the ability to defect. If they do not reach harvest max, they will spend the rest of their effort harvesting from the protected area. They will harvest up to the harvest max if possible. The resulting resources in both landscapes are multiplied by the regeneration factor, not past the carrying capacity of the landscape. Resources move between the protected and working landscapes with a probability proportional to designated probability of mobility (see Table \@ref(tab:tabAbmVars)). If agents are following a success biased learning strategy, they are paired with another agent at random and compare strategies. If one agent performs worse than another in the first time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them. If agents are following a conformist strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group. This process repeats for all time steps. This process is summarized conceptually in Fig \@ref(fig:conceptDiagram).

@@

In the multiple competing groups scenario, two groups with separate resource pools will follow the processes described above. However, at the end of each time step the groups will compare the average harvest rate each group. Whichever group has the highest average harvest rate for that timestep will report the average harvest times for the protected and working landscapes. A specified percent of the agents from the less successful group will subsequently adopt the average harvest strategy of the more successful group (see Table \@ref(tab:tabAbmVars)).

In both single and competing group scenarios, the dependent variables of interest are as follows: Cooperative behavior reports the average time that agents spend harvesting from the working landscape in each time step. Full defection is defined as all agents harvesting from the protected landscape for a proportion of time equal to the percent of the resource which is protected. Protected area degradation measures the percent carrying capacity of the protected landscape. Working landscape degradation measures the percent carrying capacity of the working landscape. Lastly, mean payoff to individuals measures the average payoff that agents receive at each time step.

Overall, the goal of this model is to test how environmental conditions affect cultural group selection mechanisms. Cultural group selection may lead to community adoption of sustainable norms, where solitary groups fail to adopt cooperative behaviors. With this model, we can test under what environmental conditions this is true and what the implications are for community wellbeing and environmental protection.

## Preliminary Work
I have developed a model which follows the above processes for one group with one resource pool. An interactive application which performs one run of that model can is found here [(click!)](https://mattclarknewaccount.shinyapps.io/ABM_Output_trial/).

@@

This project aims to extend the empirical study of cultural group selection to cooperation around environmental problems explicitly. Insights from cultural group selection may provide solutions to pressing environmental challenges of collective action. If validated experimentally, principles of cultural group selection may allow conservation scientists to better predict the viability of conservation initiatives.

@@

### Question: 
Does cultural group selection theory accurately predict real world Community Forest Management Agreements in forest adjacent wards on Pemba Island, Tanzania?

@@

* P2: If significant cultural differentiation between wards is not present in the population, wards will not adopt Community Forest Management Agreements without external influences such as nonprofit conservation initiatives.

## Research Design
Under cultural group selection theory, clear group boundaries must be present for selection to act at the group level. These group level selection pressures are what we hypothesize to drive the adoption of Community Forest Management Agreements (CoFMAs). Logically, groups with more clear differences from surrounding groups should be more likely to adopt CoFMAs than groups whose boundaries with surrounding groups are blurred. For this reason, I will calculate the cultural $F_{ST}$ values for all combinations of neighboring Pemban wards as opposed to pairwise calculations of cultural $F_{ST}$ across all wards overall.

@@

Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity. Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008]. This research will directly measure the impact of geographic and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.

@@

### Question
What effect do geographic distance and social relatedness have on transmission fidelity? Secondly, what is the relationship between transmission fidelity and behavioral outcomes?

### Predictions
* P1: Transmission fidelity will increase with increased access to behavioral demonstrators through geographic proximity and/or social connectedness.

* P2: As access to behavioral demonstration decreases, the replication fidelity of non-functional stove traits will decrease strongly relative to functional traits.

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi (aka Before Ningeze Forest) wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

@@

The survey instrument used for this research will elicit information regarding the relationship between the builder of each cookstove and the individual they copied their stove design from. Specifically, this survey tool will be used to gather data on geographic distance and communication frequency. Geographic distance will be measured according to the shortest walking distance between the two stoves. Communication frequency will be measured in terms of weekly communication and taken as a measure of social connectedness between the original stove builder and the copying individual. I will also use this survey to collect data on the decision making processes individuals exhibit when deciding to build a fuel-efficient cookstove.

To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions]. Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test. Fuel efficiency will be calculated using the following thermal efficiency equation:

$$n_{th} = \frac{C_{p}*m_{wi}*(T_{f}-T_{i})+H_{v}*(m_{wi}-m_{wf})}{(F_{i}-F_{f})(1-\frac{M}{100})*LHV_{wood}-(F_{i}-F_{f})*\frac{M}{100}*(C_{p}*(T_{b}-T_{a})+H_{v})-LHV_{char}*C_{c}} $$

where $C_{p}$ represents the heat capacity of water (4.186 J/g-K), $m_{wi}$ is the initial mass of water used, $m_{wf}$ is the mass or remaining water, $T_{i}$ and $T_{f}$ are the initial and final water temperatures respectively, and $H_{v}$ represents the enthalpy of vaporization of water (2260 J/g). $F_{i}$ and $F_{f}$ represent the initial and final masses of fuel used. $M$ is the percent moisture content of the fuelwood used, $LHV_{wood}$ and $LHV_{char}$ are the lower heating values of wood and charcoal respectively, $T_{a}$ is the ambient temperature, $T_{b}$ is the local boiling temperature of water, and $C_{c}$ is the mass of the charcoal created during the test [@defoort2009stove].

While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics. I will categorize these characteristics as functional (stove size) or nonfunctional (decorative qualities). I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from in both functional and nonfunctional characteristics.

The key research questions for this project will be assessed using multilevel regression operationalized in a Bayesian context. I will compare standardized parameter estimates, to determine which factors (geographic distance or social relatedness) most strongly contribute to copying fidelity in this study system. Similar analyses will be used to determine which factors affect cookstove primary outcome variables (fuel efficiency and particulate matter production).

@@

## Preliminary Work
In preparation for a cancelled field season in the Summer of 2020, I created a suite of forms and documents to facilitate data collection. These include:

@@

This dissertation will achieve this goal in three progressive steps. The first will formalize theory around cultural group selection in conservation contexts. This step is critical as conservation is a fundamentally different system from where these theories were originally developed. The second step will empirically validate these theoretical insights by comparing calculated cultural group selection pressures to real-world adoption of Community Forest Management Agreements on Pemba Island, Tanzania. Lastly, this project will make use of a conservation initiative currently spreading through Pemban communities, to specifically examine how behavioral efficiency may be altered by social contexts (e.g. communication frequency) as a behavior spreads throughout a population. All together, this research will allow us to better understand and possibly predict the emergence and effective spread of conservation norms within a population.
